\section{BAROSTAT}
\label{BAROSTAT}
Simulates an external pressure bath to keep the pressure or stress at the external values.
\paragraph{Attributes}
 \begin{itemize}
\item {\bf kind}:
 The type of barostat. 'Rigid' gives a barostat that keeps the internal pressure constant by allowing cell volume changes, whereas flexible allows the shape of the cell to fluctuate too.
{\\ \bf DEFAULT: }'rigid'
{\\ \bf OPTIONS: }'rigid', 'flexible'
{\\ \bf DATA TYPE: }str
\end{itemize}
 
\paragraph{Fields}
 \begin{itemize}
\item {\bf \hyperref[THERMOSTATS]{thermostat} }:
 The thermostat for the cell. Keeps the cell velocity distribution at the correct temperature.
\paragraph{Attributes}
 \begin{itemize}
\item {\bf kind}:
 The style of thermostatting. 'langevin' specifies a white noise langevin equation to be attached to the cartesian representation of the momenta. 'svr' attaches a velocity rescaling thermostat to the cartesian representation of the momenta. Both 'pile\_l' and 'pile\_g' attaches a white noise langevin thermostat to the normal mode representation, with 'pile\_l' attaching a local langevin thermostat to the centroid mode and 'pile\_g' instead attaching a global velocity rescaling thermostat. 'gle' attaches a coloured noise langevin thermostat to the cartesian representation of the momenta, and 'nm\_gle' attaches a coloured noise langevin thermostat to the normal mode representation of the momenta.
{\\ \bf DEFAULT: }'langevin'
{\\ \bf OPTIONS: }'langevin', 'svr', 'pile\_l', 'pile\_g', 'gle', 'nm\_gle', 'nm\_gle\_g'
{\\ \bf DATA TYPE: }str
\end{itemize}
 
\end{itemize}
 
