\section{ENSEMBLE}
\label{ENSEMBLE}
Holds all the information that is ensemble specific, such as the temperature and the external pressure, and the thermostats and barostats that control it.
\paragraph{Attributes}
 \begin{itemize}
\item {\bf type}:
 The ensemble that will be sampled during the simulation.
{\\ \bf DEFAULT: }'nve'
{\\ \bf OPTIONS: }'nve', 'nvt', 'npt', 'nst'
{\\ \bf DATA TYPE: }str
\end{itemize}
 
\paragraph{Fields}
 \begin{itemize}
\item {\bf stress}:
 The external stress.
{\\ \bf DIMENSION: }pressure
{\\ \bf DEFAULT: }
      [[ 1.  0.  0.]
       [ 0.  1.  0.]
       [ 0.  0.  1.]]
{\\ \bf DATA TYPE: }float
\paragraph{Attributes}
 \begin{itemize}
\item {\bf shape}:
 The shape of the array.
{\\ \bf DEFAULT: }(0,)
{\\ \bf DATA TYPE: }tuple
\end{itemize}
 
\item {\bf \hyperref[BAROSTAT]{barostat} }:
 Simulates an external pressure bath to keep the pressure or stress at the external values.
\paragraph{Attributes}
 \begin{itemize}
\item {\bf kind}:
 The type of barostat. 'Rigid' gives a barostat that keeps the internal pressure constant by allowing cell volume changes, whereas flexible allows the shape of the cell to fluctuate too.
{\\ \bf DEFAULT: }'rigid'
{\\ \bf OPTIONS: }'rigid', 'flexible'
{\\ \bf DATA TYPE: }str
\end{itemize}
 
\item {\bf \hyperref[THERMOSTATS]{thermostat} }:
 The thermostat for the atoms, keeps the atom velocity distribution at the correct temperature.
\paragraph{Attributes}
 \begin{itemize}
\item {\bf kind}:
 The style of thermostatting. 'langevin' specifies a white noise langevin equation to be attached to the cartesian representation of the momenta. 'svr' attaches a velocity rescaling thermostat to the cartesian representation of the momenta. Both 'pile\_l' and 'pile\_g' attaches a white noise langevin thermostat to the normal mode representation, with 'pile\_l' attaching a local langevin thermostat to the centroid mode and 'pile\_g' instead attaching a global velocity rescaling thermostat. 'gle' attaches a coloured noise langevin thermostat to the cartesian representation of the momenta, and 'nm\_gle' attaches a coloured noise langevin thermostat to the normal mode representation of the momenta.
{\\ \bf DEFAULT: }'langevin'
{\\ \bf OPTIONS: }'langevin', 'svr', 'pile\_l', 'pile\_g', 'gle', 'nm\_gle', 'nm\_gle\_g'
{\\ \bf DATA TYPE: }str
\end{itemize}
 
\item {\bf timestep}:
 The time step.
{\\ \bf DIMENSION: }time
{\\ \bf DEFAULT: }'1.0'
{\\ \bf DATA TYPE: }float
\item {\bf pressure}:
 The external pressure.
{\\ \bf DIMENSION: }pressure
{\\ \bf DEFAULT: }'1.0'
{\\ \bf DATA TYPE: }float
\item {\bf fixcom}:
 This describes whether the centre of mass of the particles is fixed.
{\\ \bf DEFAULT: }False
{\\ \bf DATA TYPE: }bool
\item {\bf temperature}:
 The temperature of the system.
{\\ \bf DIMENSION: }temperature
{\\ \bf DEFAULT: }1.0
{\\ \bf DATA TYPE: }float
\end{itemize}
 
